Global well-being and mental health in the internet age

### 2022-09-28 ms.Rmd
Update links
Prune pkgs

  This research was supported by the Huo Family Foundation and Economic and Social Research Council Grant ES/T008709/1. All code supporting this manuscript is available at <https://github.com/digital-wellbeing/global-wbmh>. This draft is available at psyarxiv.

abstract: | 
  In the last two decades the rise of internet technologies have impacted nearly all domains of human life and inspired concern that their widespread adoption has had a harmful effect on mental health and psychological well-being. However, research on the topic is contested and hampered by methodological shortcomings leaving the broader consequences of internet adoption unknown. We show that the past two decades have seen only small and inconsistent changes in global well-being and mental health. We demonstrate that the broadest available health and internet technology data does not support the idea that the adoption of internet and mobile broadband has been psychologically harmful on a global level. Moreover, little or no negative associations were observed for adolescents—a group thought particularly vulnerable. Further investigation of this topic requires transparent study of online behaviours where they occur: on online communication, media, social media, and video game platforms. We place an urgent call to increase the collaborative efforts between independent scientists and the internet technology sector.
keywords: well-being, mental health, internet technology, technology effects
# wordcount: 
bibliography: "references.bib"
floatsintext: yes
linenumbers: no
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
header-includes:
  - |
    \makeatletter
    \renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-1em}%
      {\normalfont\normalsize\bfseries\typesectitle}}

    \renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
      {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
      {-\z@\relax}%
      {\normalfont\normalsize\bfseries\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
    \makeatother

csl: "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass: apa7
classoption: jou
output: 
  papaja::apa6_pdf:
    number_sections: false
    keep_tex: false
  papaja::apa6_docx: 
    number_sections: false
editor_options: 
  chunk_output_type: console
always_allow_html: true
---

```{r setup, include = FALSE}
# Packages
library(papaja)
library(readxl)
library(imputeTS)
library(naniar)
library(scales)
library(janitor)
library(haven)
library(ragg)
library(khroma)
library(labelled)
library(countrycode)
library(plotly)
library(lubridate)
library(kableExtra)
library(brms)
library(knitr)
library(here)
library(posterior)
library(bayestestR)
library(gtsummary)
library(kableExtra)
library(memoise)
library(cachem)
library(ggh4x)
library(cmdstanr)
library(ggstance)
library(tidybayes)
library(ggdist)
library(DT)
library(htmlwidgets)
library(tidyverse)

opts_chunk$set(
  eval = TRUE,
  cache = TRUE,
  include = FALSE,
  warning = FALSE,
  dpi = 300,
  fig.align = "center"
)
# For tables: Show NAs as blanks
options(knitr.kable.NA = '')

# Data frame for computing over models
oo <- c(
  "Life_satisfaction", "Negative_experiences", "Positive_experiences",
  "Anxiety", "Depression", "Selfharm"
)
fits <- tibble(outcome = oo) %>%
  mutate(outcome = factor(outcome, levels = oo))

# Create directory for intermediate and output files
dir.create(here("data"), FALSE)
dir.create(here("Output"), FALSE)

# Cache for memoised functions
cd <- cache_disk("cache", max_size = Inf)

# Plotting options
# If you've defined a custom font, we will use it in outputs
theme_set(
  theme_linedraw(
    base_size = 8
  ) +
    theme(
      panel.grid = element_blank(),
      strip.text = element_text(margin = margin(4, 4, 4, 4, "pt")),
      axis.text = element_text(size = rel(0.6)),
      axis.ticks = element_line(size = rel(0.3))
    )
)
```

>>>>

  This research was supported by the Huo Family Foundation and Economic and Social Research Council Grant ES/T008709/1. All code supporting this manuscript is available at <https://doi.org/10.5281/zenodo.7004053>. This draft is available at <https://doi.org/10.31234/osf.io/9tbjy>.

abstract: |

  \textbf{Abstract.} In the last two decades the rise of internet technologies have impacted nearly all domains of human life and inspired concern that their widespread adoption has had a harmful effect on mental health and psychological well-being. However, research on the topic is contested and hampered by methodological shortcomings leaving the broader consequences of internet adoption unknown. We show that the past two decades have seen only small and inconsistent changes in global well-being and mental health. We demonstrate that the broadest available health and internet technology data does not support the idea that the adoption of internet and mobile broadband has been psychologically harmful on a global level. Moreover, little or no negative associations were observed for adolescents—a group thought particularly vulnerable. Further investigation of this topic requires transparent study of online behaviours where they occur: on online communication, media, social media, and video game platforms. We place an urgent call to increase the collaborative efforts between independent scientists and the internet technology sector.

  \textbf{Statement of Relevance} tbd

keywords: well-being, mental health, internet technology, technology effects
# wordcount: 
bibliography: "references.bib"
floatsintext: yes
linenumbers: no
draft: no
mask: no
figurelist: no
tablelist: no
footnotelist: no
# header-includes:

csl: "`r system.file('rmd', 'apa7.csl', package = 'papaja')`"
documentclass: apa7
classoption: jou
output: 
  papaja::apa6_pdf:
    number_sections: false
    keep_tex: false
  papaja::apa6_docx: 
    number_sections: false
  bookdown::html_document2:
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
always_allow_html: true
---

```{r setup, include = FALSE}
# Packages
library(papaja)
library(readxl)
library(imputeTS)
library(scales)
library(janitor)
library(haven)
library(labelled)
library(countrycode)
library(plotly)
library(lubridate)
library(kableExtra)
library(brms)
library(knitr)
library(here)
library(bayestestR)
library(ggh4x)
library(tidybayes)
library(ggdist)
library(DT)
library(tidyverse)

opts_chunk$set(
  eval = TRUE,
  cache = TRUE,
  include = knitr::is_html_output(),  # Show code only in HTML
  warning = FALSE,
  dpi = 300,
  fig.align = "center"
)
# For tables: Show NAs as blanks
options(knitr.kable.NA = '')

# Create directory for intermediate files
dir.create(here("data"), FALSE)

# Plotting options
theme_set(
  theme_linedraw(
    base_size = 10
  ) +
    theme(
      panel.grid = element_blank(),
      strip.text = element_text(margin = margin(4, 4, 4, 4, "pt")),
      axis.text = element_text(size = rel(0.6)),
      axis.ticks = element_line(size = rel(0.3))
    )
)
```
